---
title: "Homework 5"
author: "Molly Martorella"
date: "11/2/2019"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse)


knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 8,
  fig.height = 6,
  out.width = "90%")

options(knitr.kable.NA = "")

theme_set(theme_bw())

```


# Problem 1

```{r}

set.seed(10)

iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))

```

The following function, 'fix_iris':
1. fills in missing values with the mean of non-missing values for numeric columns.
2. fills in missing values with "virginica" for character columns.

```{r}

fix_iris <- function(column_to_fix){
  
  if (is.numeric(column_to_fix)) {
    
    col_mean = mean(column_to_fix, na.rm = TRUE)
    new_num_column = replace_na(column_to_fix, col_mean)
    
  } else if (is.character(column_to_fix)) {
    
    new_char_column = replace_na(column_to_fix, "virginica")
    
  } else if (!is.numeric(column_to_fix) | !is.character(column_to_fix)) {
    stop("Column vector must be numeric or character class.")
  }
  
}

```

Using a map statement to apply this function to the iris dataframe:

```{r}

map(iris_with_missing, fix_iris)

```

# Problem 2

This zip file contains data from a longitudinal study that included a control arm and an experimental arm. Data for each participant is included in a separate file, and file names include the subject ID and arm.

Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time:

Start with a dataframe containing all file names; the list.files function will help
Iterate over file names and read in data for each subject using purrr::map and saving the result as a new variable in the dataframe


```{r}

files <- tibble(
  file_name = list.files(path = "data/data_2", full.names = TRUE))

study <- 
  files %>% 
  mutate(
    file_data = map(file_name, read_csv)
  ) %>% 
  unnest(cols = file_data) %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    values_to = "observation"
  )


```
Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary
Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.
